---
title: "p8105_hw3_mp3745"
author: "Matthew Parker"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(nnet)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1

#### i)

Enter in data
```{r}
housing_df = tibble(
  contact = c(rep("low", 3), rep("high", 3)),
  home_type = rep(c("tower_block", "apartment", "house"), 2),
  sat_low = c(65, 130, 67, 34, 141, 130),
  sat_med = c(54, 76, 48, 47, 116, 105),
  sat_high = c(100, 111, 62, 100, 191, 104)
)
```

Table to compare satisfaction with contact level of residents
```{r}
# Table
sat_level_contact = housing_df %>% 
  select(-home_type) %>% 
  group_by(contact) %>% 
  summarize(
    sat_low = sum(sat_low),
    sat_med = sum(sat_med),
    sat_high = sum(sat_high),
    total = sum(sat_low, sat_med, sat_high),
    sat_low_perc = round((sat_low * 100 / total), 2),
    sat_med_perc = round((sat_med * 100 / total), 2),
    sat_high_perc = round((sat_high * 100 / total), 2)
  ) %>% 
  select(contact, sat_low, sat_low_perc, sat_med, sat_med_perc, sat_high, sat_high_perc, total) 

# View
sat_level_contact %>% 
  knitr::kable()
```

From the above table, we can see that they may be a slight association with degree of contact with other residents and their satisfaction. Of those with a high degree of contact with other residents `r sat_level_contact %>% filter(contact == "high") %>%  select(sat_high_perc)`% have a high level of satisfaction, whereas among those with a low degree of contact, `r sat_level_contact %>% filter(contact == "low") %>%  select(sat_high_perc)`% have a high level of satisfaction. Of those with a high degree of contact with other residents `r sat_level_contact %>% filter(contact == "high") %>%  select(sat_low_perc)`% have a low level of satisfaction, whereas among those with a low degree of contact, `r sat_level_contact %>% filter(contact == "low") %>%  select(sat_low_perc)`% have a low level of satisfaction.

Table to compare satisfaction with type of housing
```{r}
# Table
sat_level_housing = housing_df %>% 
  select(-contact) %>% 
  group_by(home_type) %>% 
  summarize(
    sat_low = sum(sat_low),
    sat_med = sum(sat_med),
    sat_high = sum(sat_high),
    total = sum(sat_low, sat_med, sat_high),
    sat_low_perc = round((sat_low * 100 / total), 2),
    sat_med_perc = round((sat_med * 100 / total), 2),
    sat_high_perc = round((sat_high * 100 / total), 2)
  ) %>% 
  select(home_type, sat_low, sat_low_perc, sat_med, sat_med_perc, sat_high, sat_high_perc, total) 

# View
sat_level_housing %>% 
  knitr::kable()
```

From the above table, we can see that they may be an association with housing type and residents satisfaction. Of those with a tower block `r sat_level_housing %>% filter(home_type == "tower_block") %>%  select(sat_high_perc)`% have a high level of satisfaction. Among those with a house, `r sat_level_housing %>% filter(home_type == "house") %>%  select(sat_high_perc)`% have a high level of satisfaction. And among those with an apartment, `r sat_level_housing %>% filter(home_type == "apartment") %>%  select(sat_high_perc)`% have a high level of satisfaction. Of those with a tower block `r sat_level_housing %>% filter(home_type == "tower_block") %>%  select(sat_low_perc)`% have a low level of satisfaction. Among those with a house, `r sat_level_housing %>% filter(home_type == "house") %>%  select(sat_low_perc)`% have a low level of satisfaction. And among those with an apartment, `r sat_level_housing %>% filter(home_type == "apartment") %>%  select(sat_low_perc)`% have a low level of satisfaction.  


## ii)

Fit nominal logistic regression model
```{r}
# Response matrix
resp = housing_df %>% 
  select(sat_low, sat_med, sat_high) %>% 
  as.matrix()

# Nominal model
housing_fit_nom = multinom(resp ~ contact + home_type, data = housing_df) 

# View summary
summary(housing_fit_nom)
```


Check goodness of fit
```{r}
pihat = predict(housing_fit_nom, type = 'probs') 
m = rowSums(housing_df[,3:5])

# Pearson residuals
res_pearson = (housing_df[,3:5] - pihat * m) / sqrt(pihat * m)  

# Generalized Pearson Chisq Stat
g_stat = sum(res_pearson^2) 

# P-value
pval = 1 - pchisq(g_stat, df = (6 - 4) * (3 - 1)) # fit is good
```

Since `r pval` > 0.05, we fail to reject our null hypothesis that the model does a good job fitting the data. 

Based on the signs of the coefficients in the above model, it appears there is a negative association between both low contact and renting a house with high satisfaction vs low satisfaction. There is a positive association between renting a tower block with high satisfaction vs low satisfaction. There is a negative association between low contact with medium satisfaction vs low satisfaction. There may be a slight positive association between renting a house and medium satisfaction vs low satisfaction. And there is a positive association between renting a tower block and medium satisfaction vs low satisfaction.


Odds ratios with 95% CI's
```{r}
or_ci = cbind(summary(housing_fit_nom)$coefficients, summary(housing_fit_nom)$standard.errors) %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  rename(
    coef_int = intercept,
    coef_contact_low = contactlow,
    coef_home_type_house = home_typehouse,
    coef_home_type_tower_block = home_typetower_block,
    std_err_int = v5,
    std_err_contact_low = v6,
    std_err_home_type_house = v7,
    std_err_home_type_tower_block = v8
  ) %>% 
  mutate(
    model = c("sat_med", "sat_high")
  ) %>% 
  select(model, coef_int:std_err_home_type_tower_block) %>% 
  mutate(
    int_lower = coef_int - std_err_int,
    int_higher = coef_int + std_err_int,
    contact_low_lower = coef_contact_low - std_err_contact_low,
    contact_low_higher = coef_contact_low + std_err_contact_low,
    home_type_house_lower = coef_home_type_house - std_err_home_type_house,
    home_type_house_higher = coef_home_type_house + std_err_home_type_house,
    home_type_tower_block_lower = coef_home_type_tower_block - std_err_home_type_tower_block,
    home_type_tower_block_higher = coef_home_type_tower_block + std_err_home_type_tower_block,
  ) %>% 
  mutate(
    exp_int = exp(coef_int),
    exp_int_lower = exp(int_lower),
    exp_int_higher = exp(int_higher),
    exp_contact_low = exp(coef_contact_low),
    exp_contact_low_lower = exp(contact_low_lower),
    exp_contact_low_higher = exp(contact_low_higher),
    exp_home_type_house = exp(coef_home_type_house),
    exp_home_type_house_lower = exp(home_type_house_lower),
    exp_home_type_house_higher = exp(home_type_house_higher),
    exp_home_type_tower_block = exp(coef_home_type_tower_block),
    exp_home_type_tower_block_lower = exp(home_type_tower_block_lower),
    exp_home_type_tower_block_higher = exp(home_type_tower_block_higher)
  ) %>% 
  select(model, exp_contact_low:exp_home_type_tower_block_higher)
```

